### 2022-11-30 intro.Rnw
some fixes

How can we know if a finding from a scientific study is really true? For
example, how can we know if the protective effect of a vaccine found in a study
is real? The answer to such questions is of highest importance to scientists and
decision makers, but typically we can never know for sure as any study result
comes with uncertainty.

% the sad truth is that we can never know for sure.

One way to come a little closer to the truth, however, is to repeat that
original study with new subjects. Such a \emph{replication study} may yield
similar results which would make us more confident about the original finding,
or it may yield conflicting results which would lower our confidence.
Replication studies are thus an essential part of the scientific process as they
provide a means for substantiating genuine research findings and refuting
research findings which occurred merely by chance. For this reason,
``successful'' replication is often a requirement, for example, for acceptance
of newly proposed scientific theories (\eg{} a new physical model) or the
implementation of policies based on scientific knowledge (\eg{} market approval
of a drug). The results from replication studies may thus have real world
consequences, such as deciding whether we receive the recommendation to get
vaccinated (a consequence that anyone who experienced the COVID-19 pandemic is
very aware of).

@@

The perceived value of replication studies, however, has changed over the past
decade. Earlier criticisms of low research standards \citep{Altman1994,
  Ioannidis2005} were backed up by empirical evidence. For instance,
pharmaceutical companies reported surprisingly low replication rates from
pre-clinical research \citep{Begley2012} followed by later studies estimating
that %\$28 billion
billions are wasted each year on flawed and non-replicable research in medicine
and the life sciences \citep{Chalmers2014, Freedman2015, Glasziou2018}.
Similarly, reports of fraud \citep{Wicherts2011} and questionable research
practices \citep{Wagenmakers2011, Simmons2011, John2012} sparked intense
discussion about the need for higher research standards in psychology and the
social sciences. These discussions eventually culminated in large-scale
replication projects conducted by huge researcher consortia in fields such as
psychology \citep{Opensc2015, Klein2014, Klein2018, Protzko2020}, economics
\citep{Camerer2016}, the social sciences \citep{Camerer2018}, experimental
philosophy \citep{Cova2018}, or cancer biology \citep{Errington2021}.

@@

In the midst of this doomsday, % , particularly from the hard-hit
% field of psychology,
various reforms were implemented to safe the reputation of science and to
prevent an escalation of the situation \citep[for an overview see
\eg{}][]{Munafo2017}. For instance, many journals adopted the ``registered
report'' format \citep{Chambers2021} in which a study proposal is peer reviewed
\emph{before} the study is conducted, and which, if accepted, gives provisional
publication acceptance regardless of the study outcome. Similarly, digital
infrastructure platforms, such as Zenodo (\url{https://zenodo.org/}) or the Open
Science Framework (\url{https://osf.io/}), were created to facilitate
preregistration, preprints, and code and data sharing, all of which have
substantially increased over the last decade \citep{Kidwell2016, Nosek2018,
  Rawlinson2019}. The practice of conducting replication studies has also gained
popularity, and several academic journals and funders are now explicitly
promoting and funding replication research \citep{NWO2016, NSL2018, Nature2022}.
Figure~\ref{fig:scholar} illustrates this trend via the yearly number of results
for ``\Sexpr{searchTerm}'' on Google Scholar over the last three decades. We see
that the numbers have been rapidly growing, especially after the mid 2000s (with
a minor drop in 2021, perhaps because of research slowing down due to the
COVID-19 pandemic).

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and $p$-values
    are based on the normality assumption.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)$ confidence interval &
    $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    % $[\that_{o} - \z{\alpha/2} \sigma_{o}, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} - \z{\alpha/2} \sigma_{r}, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in researcher articles, and if one of them
is missing they typically can, under some assumptions, be calculated from the
others. Using summary statistics is also often the only possible way for
conducting the replication success analysis as the raw data from the original
study may not be available to the replicators. The most important statistic is
the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time to event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations, or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate. Under the assumption of (asymptotic) normality, confidence intervals
for $\theta$ and $p$-values for testing the null hypothesis
$H_{0} \colon \theta = 0$ can be computed as shown in Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not required to show the same direction as this is taken into account by the
$p$-values. Some replication projects \citep{Opensc2015, Errington2021} also
defined replication success via simultaneous non-significance of original and
replication study ($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this
definition ``replication success'' can almost always be achieved by conducting
original and replication study with very few samples so that the $p$-values are
large. The approach is also logically questionable as it could be seen as an
instance of the ``absence of evidence fallacy'' \citep{Altman1995} meaning the
failure to find evidence against the null hypothesis is erroneously interpreted
as evidence for the null hypothesis. Meta-analytic extensions of the
significance approach define replication success by significance of a combined
effect estimate.\footnote{The meta-analytic approach can also be given a
  Bayesian interpretation: A flat prior distribution for the underlying effect
  size $\theta$ is updated by the data from original and replication study.
  Replication success defined by a meta-analytic $p$-value being smaller than
  $\alpha$ is then equivalent to replication success via a Bayesian posterior
  probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
  respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$
  depending on the direction of combined estimate.} Typically, the assumption of
a common underlying effect size is seen as reasonable so that fixed-effects
meta-analysis is used for pooling. Random-effects meta-analysis has primarily
been used if more than one replication study of the same original study are
conducted since in this case replicators are often interested in also
understanding between-replication heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

\subsubsection{Example: Reproducibility Project Cancer Biology}
I will now illustrate the assessment of replicability on data from the
Reproducibility Project: Cancer Biology \citep{Errington2021}. This large-scale
project attempted to replicate 53 landmark studies from the field of cancer
biology. However, due to various difficulties (\eg{} missing information from
the original studies or problems in conducting the experiments) only 23 of the
53 studies could be repeated. \citet{Errington2021} report that these
experiments led to data on 158 quantitative effects. However, from the data
which they provide only 132 quantitative effects are provided with original and
replication standardized mean difference effect estimates along with standard
errors, and only these data will be used in the subsequent analyses.

@@

Figure~\ref{fig:upset} shows how many of the replications are successful
according to the criteria from Table~\ref{tab:rsdef}, and their combinations.
For the total 132 replications, most successes occur for the meta-analytic
significance (\Sexpr{nma}) and the prediction interval criteria (\Sexpr{npi}),
followed by significance (\Sexpr{nsig}), and relative effect size
(\Sexpr{nres}). However, looking at the combinations of the criteria, only in
two replications are all of them satisfied simultaneously (green).

@@

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the Reproducibility Project: Cancer Biology
    \citep{Errington2021}. (P, X, E) denotes effect number E from experiment
    number X, from original paper number P. Shown are original $\that_{o}$,
    replication $\that_{r}$, and pooled effect estimate $\that_{m}$ with 95\%
    confidence intervals, variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$,
    relative effect estimate $d = \that_{r}/\that_{o}$, original $p$-value
    $p_{o}$, replication $p$-value $p_{r}$, meta-analytic $p$-value $p_{m}$,
    $Q$-test $p$-value $p_{Q}$, and sceptical $p$-value $p_{\mathrm{S}}$. A
    level $\alpha = 5\%$ is used as threshold for replication success via
    $p$-values, and a threshold $d_{\text{min}} = 1$ for the relative effect
    estimate.}
  \label{tab:RPCB}
% \resizebox{1\textwidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## select a subset with one effect per paper
set.seed(7)
rpcbTable <- rpcb |>
    na.omit() |>
    filter(smdo < 80, smdr < 80) |>
    group_by(paper) |>
    sample_n(size = 1) |>
    ungroup() |>
    arrange(pr2) |>
    select(id, smdoFormat, ## so,
           smdrFormat, smdmFormat,
           c, dFormat, ## sr,
           po2Format, pr2Format, pm2Format, pQFormat, ps2Format
           )

@@

Bayesian inference is an approach to statistical inference where Bayes' theorem
is used to make probability statements about unknown parameters based on the
observed data. The central quantity for doing so is the distribution of the
parameters conditional on the data, called the \emph{posterior distribution}. It
can be obtained from Bayes' theorem
\begin{align*}
  f(\theta \given \text{data})
  = f(\theta) \times \frac{f(\text{data} \given \theta)}{f(\text{data})}
\end{align*}
meaning that the \emph{prior distribution} for the parameter $\theta$ with
density/probability mass function $f(\theta)$ is multiplied by the (normalized)
likelihood of the data, also know as \emph{Bayesian updating}. Parameter values
which increase the likelihood of the data become more likely \emph{a posteriori}
but they are also weighted by their \emph{a priori} plausibility through the
prior. As such, Bayesian inference provides a formal way for combining
information from the data at hand with external knowledge encoded in the prior.

Many consider the prior to be also the weak point of Bayesian inference since it
is unclear how it should be specified in the absence of external knowledge. The
\emph{reverse-Bayes} approach, first proposed by \citet{Good1950}, is one way of
dealing with this issue. The idea is to reverse Bayes' theorem
\begin{align*}
  f(\theta) = f(\theta \given \text{data}) \times
  \frac{f(\text{data})}{f(\text{data} \given \theta)}
\end{align*}
and instead ``downdate'' a posterior with the observed data. So, in contrast to
the conventional ``forward-Bayes'' approach where we start with a prior, update
it with the data, and end up with a posterior, the reverse-Bayes approach starts
with the posterior and ends up with the prior. Reverse-Bayes inference then
revolves around the question of whether the resulting prior is plausible in
light of external knowledge, and if so, this could be seen as support for the
specified posterior.

To illustrate reverse-Bayes inference, let us return to the replication setting.
Assume we want to conduct inference about the unknown effect size $\theta$ based
on the effect estimate from the original study $\that_{o}$ and its standard
error $\sigma_{o}$. We will assume that $\that_{o}$ is normally distributed
around the unknown effect size $\theta$ with (known) variance equal to its
squared standard error $\sigma^{2}_{o}$, here and henceforth denoted by
$\that_{o} \given \theta \sim \Nor(\theta, \sigma^{2}_{o})$. Furthermore, we
specify a zero-mean normal prior with variance $\tau^{2}$ for the effect size
$\theta$, representing the position of a sceptic who does not believe in the
presence of a non-zero effect. The ``stubbornness'' of the sceptic is determined
by how large the variance $\tau^{2}$ is chosen. Combining the sceptical prior
with the likelihood produces then a posterior which is again normal
$\theta \given \that_{o}, \sigma_{o} \sim \Nor(\mu_{\text{post}}, \sigma^{2}_{\text{post}})$
with mean and variance
\begin{align*}
  &\mu_{\text{post}} = \frac{\that_{o}}{1 + \sigma^{2}_{o}/\tau^{2}}&
  &\text{and}&
  &\sigma^{2}_{\text{post}} = \frac{1}{1/\sigma^{2}_{o} + 1/\tau^{2}}.&
\end{align*}
The associated $(1- \alpha)$ highest posterior density credible interval is
given by
\begin{align}
  \left[\mu_{\text{post}} \pm z_{\alpha/2} \, \sigma_{\text{post}}\right]
  \label{eq:postCI}
\end{align}
and if this credible interval excludes parameter values smaller/larger than zero
(depending on the orientation of the effect size) this may be interpreted as
evidence\footnote{For readers who do not agree with this notion of evidence, do
  not panic. In this thesis (Paper I and III) we extend this reverse-Bayes
  procedure to use alternative measures of evidence, such as Bayes factors.} for
a genuine effect found in the original study.

@@

ggplot(data = plotDF, aes(x = type, y = t)) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.3) +
    geom_pointrange(aes(ymin = t - za*s, ymax = t + za*s, col = type),
                    fatten = 3,
                    position = position_dodge2(width = dodgewidth),
                    show.legend = FALSE) +
    annotate("segment", x = 1.85, y = -0.5, xend = 1.98, yend = -0.05,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 1.85, y = -0.6, size = 3, alpha = 0.6,
             label = paste0(round(100*alpha/2, 2), "% quantile fixed at zero")) +
    annotate("segment", x = 1.6, y = 2.25, xend = 2.7, yend = 2.25,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 2.1, y = 2.45, size = 3, alpha = 0.6,
             label = "reverse-Bayes analysis") +
    annotate("segment", x = 3, y = 2.25, xend = 4, yend = 2.25,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 3.45, y = 2.45, size = 3, alpha = 0.6,
             label = "assess prior-data conflict") +
    annotate("path", x = 3 + dpriorScaled, y = xprior, color = 2, alpha = 0.5) +
    annotate("path", x = 2 + dpostScaled, y = xpost, color = 4, alpha = 0.5) +
    geom_text(aes(label = Label, y = t - za*s*0.9),
              position = position_dodge2(width = dodgewidth),
              parse = TRUE, size = 2.75, alpha = 0.8, hjust = -0.08) +
    labs(x = NULL, y = "Effect size") +
    scale_color_manual(values = c("Original" = 1, "Replication" = 1,
                                  "Posterior" = 4, "Prior" = 2)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.x = element_text(size = 11))
@
\caption{Illustration of reverse-Bayes assessment of replication success using
  data from the original study (Paper 9, Experiment 2, Effect 5) and its three
  replication studies from the Reproducibility Project: Cancer Biology
  \citep{Errington2021}. Shown are effect estimates and prior/posterior means
  with \Sexpr{100*(1-alpha)}\% confidence/credible interval. The original finding is
  challenged with a ``sceptical'' prior, sufficiently concentrated around zero so
  that the resulting posterior is no longer convincing at level
  $\alpha = \Sexpr{100*(alpha)}\%$. Prior predictive $p$-values $p_{\text{Box}}$ are
  computed for quantifying prior-data conflict.}
\label{fig:RBillustration}
\end{figure}

Figure~\ref{fig:RBillustration} illustrates the derivation of the sufficiently
sceptical prior variance for an original study included in the Reproducibility
Project: Cancer Biology with standardized mean difference effect estimate
$\that_{o} = \Sexpr{round(to, 2)}$ and standard error
$\sigma_{o} = \Sexpr{round(so, 2)}$. We see that a sufficiently sceptical prior
with variance $\tau_{\alpha}^{2} = \Sexpr{round(sprior, 2)}^{2}$ is required to
render the resulting posterior no longer convincing at level
$\alpha = \Sexpr{round(100*alpha, 2)}$\%.

>>>>

How can we know if a finding from a scientific study is really true? For
example, how can we know if the protective effect of a vaccine found in a study
is real? The answer to such questions is of high importance to scientists and
decision makers. However, because any study result comes with uncertainty we can
never know for sure.

One way to come a little closer to the truth is to repeat the original study
with new subjects. Such a \emph{replication study} may yield similar results
which would make us more confident about the original finding, or may yield
conflicting results which would lower our confidence. Replication studies are
thus an essential part of the scientific process as they provide a means for
substantiating genuine research findings and refuting research findings which
occurred merely by chance. For this reason, ``successful'' replication is often
a requirement, for example, for acceptance of newly proposed scientific theories
(\eg{} a new physical model) or the implementation of policies based on
scientific knowledge (\eg{} market approval of a drug). The results from
replication studies may thus have real world consequences, such as deciding
whether we receive the recommendation to get vaccinated (a consequence that
anyone who experienced the COVID-19 pandemic is very aware of).

@@

The perceived value of replication studies has changed over the past decade.
Earlier criticisms of low research standards \citep{Altman1994, Ioannidis2005}
were backed up by empirical evidence. For instance, pharmaceutical companies
reported surprisingly low replication rates from pre-clinical research
\citep{Begley2012} followed by later studies estimating that %\$28 billion
billions are wasted each year on flawed and non-replicable research in medicine
and the life sciences \citep{Chalmers2014, Freedman2015, Glasziou2018}.
Similarly, reports of fraud \citep{Wicherts2011} and questionable research
practices \citep{Wagenmakers2011, Simmons2011, John2012} sparked intense
discussion about the need for better research standards in psychology and the
social sciences. These discussions eventually culminated in large-scale
replication projects conducted by huge researcher consortia in fields such as
psychology \citep{Opensc2015, Klein2014, Klein2018, Protzko2020}, economics
\citep{Camerer2016}, the social sciences \citep{Camerer2018}, experimental
philosophy \citep{Cova2018}, and cancer biology \citep{Errington2021}.

@@

In the midst of this doomsday, % , particularly from the hard-hit
% field of psychology,
various reforms were implemented to prevent an escalation of the situation and
save the reputation of science \citep[for an overview see \eg{}][]{Munafo2017}.
For instance, many journals adopted the ``registered report'' format
\citep{Chambers2021} in which a study proposal is peer reviewed \emph{before}
the study is conducted, and which, if accepted, gives provisional publication
acceptance regardless of the study outcome. Similarly, digital infrastructure
platforms, such as Zenodo (\url{https://zenodo.org/}) or the Open Science
Framework (\url{https://osf.io/}), were created to facilitate preregistration,
preprints, code and data sharing, all of which have substantially increased over
the last decade \citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of
conducting replication studies has also gained popularity. Several academic
journals and funders are now explicitly promoting replication research
\citep{NWO2016, NSL2018, Nature2022}. Figure~\ref{fig:scholar} illustrates this
trend via the yearly number of results for the search term
``\Sexpr{searchTerm}'' on Google Scholar over the last three decades. We see
that the numbers have been rapidly growing, especially after the mid 2000s, with
a minor drop in 2021 (perhaps because of research slowing down due to the
COVID-19 pandemic).

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values are based on the normality assumption.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)$ confidence interval &
    $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    % $[\that_{o} - \z{\alpha/2} \sigma_{o}, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} - \z{\alpha/2} \sigma_{r}, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in researcher articles, and if one of them
is missing they typically can, under some assumptions, be calculated from the
others. Using summary statistics is also often the only possible way for
conducting the replication success analysis as the raw data from the original
study may not be available to the replicators. The most important statistic is
the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time to event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate. Under the assumption of (asymptotic) normality, confidence intervals
for $\theta$ and $p$-values for testing the null hypothesis
$H_{0} \colon \theta = 0$ can be computed as shown in Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not required to show the same direction as this is taken into account by the
$p$-values. Some replication projects \citep{Opensc2015, Errington2021} also
defined replication success via simultaneous non-significance of original and
replication study ($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this
definition ``replication success'' can almost always be achieved by conducting
original and replication study with very few samples so that the $p$-values are
large. The approach is also logically questionable as it could be seen as an
instance of the ``absence of evidence fallacy'' \citep{Altman1995} meaning the
failure to find evidence against the null hypothesis is erroneously interpreted
as evidence for the null hypothesis. Meta-analytic extensions of the
significance approach define replication success by significance of a combined
effect estimate.\footnote{The meta-analytic approach can also be given a
  Bayesian interpretation: A flat prior distribution for the underlying effect
  size $\theta$ is updated by the data from original and replication study.
  Replication success defined by a meta-analytic $p$-value being smaller than
  $\alpha$ is then equivalent to replication success via a Bayesian posterior
  probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
  respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$
  depending on the direction of combined estimate.} Typically, the assumption of
a common underlying effect size is seen as reasonable so that fixed-effects
meta-analysis is used for pooling. Random-effects meta-analysis has primarily
been used if more than one replication study of the same original study are
conducted since replicators are often interested in understanding
between-replication heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

\subsubsection{Example: Reproducibility Project Cancer Biology}
I will now illustrate the assessment of replicability on data from the
Reproducibility Project: Cancer Biology \citep{Errington2021}. This large-scale
project attempted to replicate 53 landmark studies from the field of cancer
biology. However, due to various difficulties (\eg{} missing information from
the original studies or problems in conducting the experiments) only 23 of the
53 studies could be repeated. \citet{Errington2021} report that these
experiments led to data on 158 quantitative effects. However, from the data
which they provide only 132 quantitative effects are provided with original and
replication standardized mean difference effect estimates along with standard
errors, and only they will be used in the subsequent analyses.

@@

Figure~\ref{fig:upset} shows how many of the replications are successful
according to the criteria from Table~\ref{tab:rsdef} (except the confidence
interval criteria since they do not take into account the uncertainty from both
studies). For the total 132 replications, most successes occur for the
meta-analytic significance (\Sexpr{nma}) and the prediction interval criteria
(\Sexpr{npi}), followed by significance (\Sexpr{nsig}), and relative effect size
(\Sexpr{nres}). However, looking at the combinations of the criteria, only in
two replications are all criteria satisfied simultaneously (green).

@@

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the Reproducibility Project: Cancer Biology
    \citep{Errington2021}. (P, X, E) denotes effect number E from experiment
    number X, from original paper number P. Shown are original $\that_{o}$,
    replication $\that_{r}$, and pooled effect estimate $\that_{m}$ with 95\%
    confidence intervals, variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$,
    relative effect estimate $d = \that_{r}/\that_{o}$, original
    \textit{p}-value $p_{o}$, replication \textit{p}-value $p_{r}$,
    meta-analytic \textit{p}-value $p_{m}$, $Q$-test \textit{p}-value $p_{Q}$,
    and sceptical \textit{p}-value $p_{\mathrm{S}}$. A level $\alpha = 5\%$ is
    used as threshold for replication success via \textit{p}-values, and a
    threshold $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
% \resizebox{1\textwidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## select a subset with one effect per paper
set.seed(7)
rpcbTable <- rpcb |>
    na.omit() |>
    filter(smdo < 80, smdr < 80) |>
    group_by(paper) |>
    sample_n(size = 1) |>
    ungroup() |>
    arrange(pr2) |>
    select(id, smdoFormat, ## so,
           smdrFormat, smdmFormat,
           c, dFormat, ## sr,
           po2Format, pr2Format, pm2Format, pQFormat, ps2Format
           )

@@

Bayesian inference is an approach to statistical inference where Bayes' theorem
is used to make probability statements about unknown parameters based on the
observed data. The central quantity for doing so is the distribution of the
parameters conditional on the data, called the \emph{posterior distribution}. It
can be obtained from Bayes' theorem
\begin{align*}
  f(\theta \given \text{data})
  = f(\theta) \times \frac{f(\text{data} \given \theta)}{f(\text{data})}
\end{align*}
meaning that the \emph{prior distribution} for the parameter $\theta$ with
probability density/mass function $f(\theta)$ is multiplied by the (normalized)
likelihood of the data, also know as \emph{Bayesian updating}. Parameter values
which increase the likelihood of the data become more likely \emph{a posteriori}
but they are also weighted by their \emph{a priori} plausibility through the
prior. As such, Bayesian inference provides a formal way for combining
information from the data at hand with external knowledge encoded in the prior.

Many consider the prior to be also the weak point of Bayesian inference since it
is unclear how it should be specified in the absence of external knowledge. The
\emph{reverse-Bayes} approach, first proposed by \citet{Good1950}, is one way of
dealing with this issue. The idea is to reverse Bayes' theorem
\begin{align*}
  f(\theta) = f(\theta \given \text{data}) \times
  \frac{f(\text{data})}{f(\text{data} \given \theta)}
\end{align*}
and instead ``downdate'' a posterior with the observed data. So, in contrast to
the conventional ``forward-Bayes'' approach where we start with a prior, update
it with the data, and end up with a posterior, the reverse-Bayes approach starts
with the posterior and ends up with the prior. For example, the posterior may be
specified to indicate evidence for/against a particular hypothesis under
investigation. Reverse-Bayes inference then revolves around the question of
whether the resulting prior is plausible in light of external knowledge, and if
so, this could be seen as support for the specified posterior.

To illustrate reverse-Bayes inference, let us return to the replication setting.
Assume we want to conduct inference about the unknown effect size $\theta$ based
on the effect estimate from the original study $\that_{o}$ and its standard
error $\sigma_{o}$. We will assume that $\that_{o}$ is normally distributed
around the unknown effect size $\theta$ with (known) variance equal to its
squared standard error $\sigma^{2}_{o}$, here and henceforth denoted by
$\that_{o} \given \theta \sim \Nor(\theta, \sigma^{2}_{o})$. Furthermore, we
specify a zero-mean normal prior with variance $\tau^{2}$ for the effect size
$\theta$, representing the position of a sceptic who does not believe in the
presence of a genuine effect. The ``stubbornness'' of the sceptic is determined
by how large the variance $\tau^{2}$ is chosen. Combining the sceptical prior
with the likelihood produces a posterior which is again normal
$\theta \given \that_{o}, \sigma_{o} \sim \Nor(\mu_{\text{post}}, \sigma^{2}_{\text{post}})$
with mean and variance
\begin{align*}
  &\mu_{\text{post}} = \frac{\that_{o}}{1 + \sigma^{2}_{o}/\tau^{2}}&
  &\text{and}&
  &\sigma^{2}_{\text{post}} = \frac{1}{1/\sigma^{2}_{o} + 1/\tau^{2}}.&
\end{align*}
The associated $(1- \alpha)$ highest posterior density credible interval is
given by
\begin{align}
  \left[\mu_{\text{post}} \pm z_{\alpha/2} \, \sigma_{\text{post}}\right]
  \label{eq:postCI}
\end{align}
and if this credible interval excludes parameter values smaller/larger than zero
(depending on the orientation of the effect size) this may be interpreted as
evidence\footnote{For readers who do not agree with this notion of evidence, do
  not panic. In this thesis (Paper I and IV) we extend this reverse-Bayes
  procedure to use alternative measures of evidence, such as Bayes factors.} for
a genuine effect found in the original study.

@@

ggplot(data = plotDF, aes(x = type, y = t)) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.3) +
    geom_pointrange(aes(ymin = t - za*s, ymax = t + za*s, col = type),
                    fatten = 3,
                    position = position_dodge2(width = dodgewidth),
                    show.legend = FALSE) +
    annotate("segment", x = 1.85, y = -0.5, xend = 1.98, yend = -0.05,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 1.85, y = -0.6, size = 3, alpha = 0.6,
             label = paste0(round(100*alpha/2, 2), "% quantile fixed at zero")) +
    annotate("segment", x = 1.6, y = 2.25, xend = 2.7, yend = 2.25,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 2.1, y = 2.45, size = 3, alpha = 0.6,
             label = "reverse-Bayes analysis") +
    annotate("segment", x = 3, y = 2.25, xend = 4, yend = 2.25,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 3.45, y = 2.45, size = 3, alpha = 0.6,
             label = "assess prior-data conflict") +
    annotate("path", x = 3 + dpriorScaled, y = xprior, color = 2, alpha = 0.5) +
    annotate("path", x = 2 + dpostScaled, y = xpost, color = 4, alpha = 0.5) +
    geom_text(aes(label = Label, y = t - za*s*0.9),
              position = position_dodge2(width = dodgewidth),
              parse = TRUE, size = 2.75, alpha = 0.8, hjust = -0.08) +
    labs(x = NULL, y = "Effect size") +
    scale_color_manual(values = c("Original" = 1, "Replication" = 1,
                                  "Posterior" = 4, "Prior" = 2)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.x = element_text(size = 11))
@
\caption{Illustration of reverse-Bayes assessment of replication success using
  data from the original study (Paper 9, Experiment 2, Effect 5) and its three
  replication studies from the Reproducibility Project: Cancer Biology
  \citep{Errington2021}. Shown are effect estimates and prior/posterior means
  with \Sexpr{100*(1-alpha)}\% confidence/credible interval. The original finding is
  challenged with a sceptical prior, sufficiently concentrated around zero so
  that the resulting posterior is no longer convincing at level
  $\alpha = \Sexpr{100*(alpha)}\%$. Prior predictive \textit{p}-values $p_{\text{Box}}$ are
  computed for quantifying prior-data conflict.}
\label{fig:RBillustration}
\end{figure}

Figure~\ref{fig:RBillustration} illustrates the derivation of the sufficiently
sceptical prior variance for an original study included in the Reproducibility
Project: Cancer Biology with standardized mean difference effect estimate
$\that_{o} = \Sexpr{round(to, 2)}$ and standard error
$\sigma_{o} = \Sexpr{round(so, 2)}$. We see that the sufficiently sceptical
prior with variance $\tau_{\alpha}^{2} = \Sexpr{round(sprior, 2)}^{2}$ produces
a posterior with $1 - \alpha = \Sexpr{round(100*(1 - alpha), 2)}\%$ credible
interval fixed to zero, so that the original finding is rendered no longer
convincing at level $\alpha = \Sexpr{round(100*alpha, 2)}$\%.
